Insect venom anaphylaxis: a case - control study of the European Anaphylaxis Registry

### 2019-05-23 analysis/paper/paper.Rmd
figure symptoms

@@

## Demographics

We saw a significant difference in the clinical features of insect anaphylaxis in children and adults. The younger popoulation significantly less often had concomitant conditions (i.e. DM, HT, malignant diseases and mastocytosis), and more often presented with atopic dermatitis, rhinitis and asthma. Therefore, we decided to adjust the analysis for age and sex.

```{r}
demoTab %>% rbind(demo_age_tab) %>% t() %>%  
  {data.frame(` `=rownames(.),.)} %>%
#   flextable(col_keys = c("X.","female","male","children.adolescents","adults")) 
# topology <- data.frame(col_keys = c("X.","female","male","children.adolescents","adults"),
#                        colB = c("","Sex","Sex","Age","Age"),
#                        ColA = c("","Female","Male","<18",">=18"))
# tempT %>%  
#   set_header_df(mapping = topology,key = "col_keys") %>% 
#   merge_h(part="header") %>% 
#   merge_v() %>% 
#   empty_blanks() %>% 
#   autofit()
knitr::kable(col.names = c("","female","male","<18","adults"),row.names = F)
```

```{r, results='asis'}
# st_options('bootstrap.css', FALSE)
# #rdb %>% select()
# ctable(rdb$d_centres_country %>% sapply(function(x){
#   as.character(x) %>% factor()
# }), rdb$q_340_insects%>% sapply(function(x){
#   as.character(x) %>% factor()
# }), style = 'rmarkdown',
#        useNA = "no",
#        prop = "r",
#        totals = F,
#        round.digits = 1,
# dnn = c("Countries","Insects"))

@@

```{r insectstime, fig.cap="Proportion of anaphylaxis cases elicited by specific insects according to the month in which the reaction occured. Less frequent insects were groupeped together as 'other'. The lowest panel shows the proportion of insect elicited cases to ceses elicited by other triggers.", fig.height=6,fig.width=4}
#### plot MOR#####
plot_MOR <- gridExtra::grid.arrange(
  #cowplot::plot_grid(
  rdbp %>%
  select(b_reactiondate,grouping,q_340_insects,d_centres_country) %>%
  mutate(MOR = substr(b_reactiondate,4,5)) %>%
  filter(!is.na(q_340_insects),MOR!="00") %>%
  ggplot(aes(MOR,fill=q_340_insects))+
  geom_bar(position = "fill")+
    theme_classic()+
  theme(axis.text.x = element_blank(),
        axis.title.x = element_blank(),
        axis.ticks.x = element_blank(),
        legend.position = "none"
        )+
  ylab("Proportion"),
rdbp %>%
  select(b_reactiondate,grouping,q_340_insects,d_centres_country) %>%
  mutate(MOR = substr(b_reactiondate,4,5)) %>%
  filter(!is.na(q_340_insects),MOR!="00") %>%
  ggplot(aes(MOR,fill=q_340_insects))+
  geom_bar()+
  theme_classic()+
  labs(fill="Insect",x = "Month of the year")+
  theme(axis.text.x = element_blank(),
        axis.title.x = element_blank(),
        axis.ticks.x = element_blank(),
        legend.position = c(0.01,.98),
        legend.justification = c(0,1)),
rdbp %>%
  select(b_reactiondate,grouping,d_elicitor_gr5,d_centres_country) %>%
  mutate(MOR = substr(b_reactiondate,4,5),
         d_elicitor_gr5 = relevel(d_elicitor_gr5, "insects")) %>%
  filter(!is.na(grouping),MOR!="00") %>%
  group_by(MOR,grouping) %>%
  summarize(n = n()) %>%
  group_by(MOR) %>%
  summarise(prop = n[1]/sum(n)) %>%
  ggplot(aes(MOR,prop))+
  geom_bar(stat="identity")+
  theme_classic()+
  labs(x = "Month of the year",y = "Fraction of insect elicited ANA"),
  #theme(legend.position = c(0.01,.98),
  #      legend.justification = c(0,1)),
heights = c(0.4,1,0.5),
ncol = 1
)
```

Insect venom elicited anaphylaxis in contrast to other elicitors showed a significant seasonal fluctuation and was most frequently reported from May till October. Their proportion to other cases during the summer seasons reached 60% and was as under 1% of cases during winter. Yellow-jacket was the most prominent causing-insect reported in these cases foolowed by bee (Fig \@ref(fig:insectstime)).

```{r MORaltogether}

```{r insectsgeo, fig.cap="Proportion of anaphylaxis cases elicited by specific insects in reporting countries. Less frequent insects were groupeped together as 'other'. Countries which reported less than 20 cases of anaphylaxis due to insect venom were not illustrated in this figure."}
```

The Figure \@ref(fig:insectsgeo) shows the distribution of insects according to the geographic area (countries). The most common elicitor among most of the countries were yellow-jackets. Switzerland had significantly more reactions to bees than Germany. Centers from Italy reported more cases elicited by hornets in comparison to other countries.

```{r age}
cdplot(rdb$grouping~as.numeric(rdb$d_age))
```

Cases of IVA were more requent in oder patients. Children suffered mainly form food elicited anaphylaxis. Therefore we adjusted the anlysis apropriately for age.

```{r}
mosaicplot(formula = d_severity_rm~d_insect_gr4,data = rdb)
rdb %>% filter(!is.na(d_insect_gr4)) %>% 
ggplot(aes(d_insect_gr4,fill=factor(d_severity_rm)))+
  geom_bar(position="fill")+
  facet_grid(.~d_age_gr2)
```

## Symptoms
```{r}
testInsectsbinomialTab$symptoms[,-7] %>% knitr::kable()
```

```{r}
plotSympt
```

@@

### Adrenaline use is lower in patients with IVA compared to other anaphylaxis cases

@@

Patients who underwent IVA significantly less often recived adrenaline treatment than in other adrenaline cases. After adjusting both groups for simmilar age, sex and severity distribution - the difference in adrenaline use was still significant irrespective of the route of administration. Antihistaminic drugs on the other had were given more often in cases of IVA when compared to non-IVA.

@@

```{r}
ANAscore_matched %>%
        mutate(grouping = recode(grouping,
                                 'insects'='IVA',
                                 'other'='non-IVA'),
               d_522_adren_agg = recode(d_522_adren_agg,
                                       'no'='no adrenaline',
                                       'yes'='adrenaline given') ) %>%
                         group_by(grouping,
                                  #d_AAI_prescribed,
                                  d_522_adren_agg) %>%
                         summarize(n = n()) %>%
  pull(n) %>% 
  matrix(nrow=2) %>% 
  chisq.test() %>% 
```

@@

```{R adrenalineuse, fig.cap = "The proportion of insect venom elicited anaphylaxis cases that were treated with adrenaline in comparison to other elicitors."}
F2$plot2
```

>>>>

We saw a significant difference in the clinical features of insect anaphylaxis in children and adults. The younger popoulation significantly less often had concomitant conditions (i.e. DM, HT, malignant diseases and mastocytosis), and more often presented with atopic dermatitis, rhinitis and asthma. Therefore, we decided to adjust the analysis for age and sex to reduce the comparison bias. When comparing the management of both types of anaphylaxis we also matched the control group according to severity.

@@

@@

```{r insectstime, fig.cap="A: Proportion of anaphylaxis cases elicited by specific insects according to the month in which the reaction occured. Less frequent insects were groupeped together as 'other'. B: The proportion of insect elicited cases to ceses elicited by other triggers considering patient's age. C: Geographical differences in the most common eliciotors of IVA. Countries which reported less than 10 cases of anaphylaxis due to insect venom were not illustrated in this figure."}
fig_basic
```

Insect venom elicited anaphylaxis in contrast to other elicitors showed a significant seasonal fluctuation and was most frequently reported from May till October. Their proportion to other cases during the summer seasons reached 60% and was as under 1% of cases during winter. Yellow-jacket was the most prominent IVA-causing insect reported in these cases followed by bees. The IVA causing insects differed in European countries with hornets being more prominient on the south of europe. IVA was more frequent in adults and seniors when compared to children and youg adults. (Fig \@ref(fig:insectstime)).

Cases of IVA were more frequent in older patients. Children suffered mainly from food elicited anaphylaxis. Therefore we adjusted the anlysis apropriately for age.

## Symptoms

@@

### Adrenaline is less often used in patients with IVA compared to other anaphylaxis cases

@@

Patients who underwent IVA significantly less often recived adrenaline treatment than in other adrenaline cases (`r test_adrenuse1$observed %>% prop.table()`, `r test_adrenuse1$p.value %>% pval()`). After adjusting both groups for simmilar age, sex and severity distribution - the difference in adrenaline use was still significant irrespective of the route of administration. Antihistaminic drugs on the other had were given more often in cases of IVA when compared to non-IVA.

@@

@@

```{R adrenalineuse, fig.cap = "The proportion of insect venom elicited anaphylaxis cases that were treated with adrenaline in comparison to other elicitors."}
fig_adrenuse1
```

```{r figsymptoms, fig.cap="Symptoms of insect venom anaphylaxis (IVA) compared to other elicitors",fig.width=8}
fig_symptoms
```
